SNIK Graph---Visualizing Knowledge about Management of Hospital Information Systems

### 2021-11-20 paper.tex
Note about repeated star operations as this was not understood by a reviewer, see https://github.com/snikproject/snik-graph/issues/364.
Switch to ICSCR 2022 format.

%===========[ PREAMBLE ]========================================================
\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage[a4paper, top=0.98in, bottom=0.79in, left=0.98in, right=0.98in]{geometry}
\usepackage{graphicx}
\usepackage{tabularx}
\usepackage{fancyhdr}
\usepackage{setspace}
\usepackage{titlesec}
\usepackage{enumitem}
\usepackage{todonotes}
\usepackage{csquotes}
\usepackage{listings}
\lstset{language=SPARQL,breaklines=true}
\usepackage{siunitx}
\usepackage[labelsep=period]{caption}
\usepackage{mwe}
\usepackage{xfrac}
\usepackage[colorlinks=true,urlcolor=blue]{hyperref}
\usepackage{cleveref}
\usepackage{aurl}
\daurl{bb}{http://www.snik.eu/ontology/bb/}
\daurl{ob}{http://www.snik.eu/ontology/bb/}
\daurl{meta}{http://www.snik.eu/ontology/meta/}
\graphicspath{{./img/}}
\usepackage[backend=biber, style=ieee]{biblatex} %For bibliography management with biblatex
\addbibresource{paper.bib}
\addbibresource{snik.bib}
\addbibresource{relatedwork.bib}

% ==========[ FONT ]============================================================
%For the font of the document uncoment one of the following options.
%If using pdflatex compiler the follwoing commands will provide an arial-like
%font (helvetica)
\usepackage{helvet}
\renewcommand{\familydefault}{\sfdefault}
%If using xelatex compiler the follwoing commands will provide arial font
%\usepackage{fontspec}
%\setmainfont{Arial}

%===========[ FORMAT ]==========================================================
%Please do not edit or remove these lines
\pagenumbering{gobble}
\setlength{\parskip}{6pt}
\setlist[1]{topsep=0pt,itemsep=-5pt}
\addto\captionsenglish{\renewcommand{\figurename}{Figure}}
\addto\captionsenglish{\renewcommand{\tablename}{Table}}
\titlespacing*{\section}{0pt}{10pt}{4pt}
\titlespacing*{\subsection}{0pt}{0ex}{0pt}
\titlespacing*{\subsubsection}{0pt}{0pt}{0pt}
\titlespacing*{\paragraph}{0pt}{6pt}{6pt}
\captionsetup{belowskip=0pt,skip=6pt,labelfont=bf}%justification=raggedright,singlelinecheck=false
\makeatletter
\renewcommand{\maketitle}{
    \begin{flushleft}
        \footnotesize{Placeholder Name of the conference}\\\medskip
        \footnotesize{Placeholder Session title}\\\medskip
        \footnotesize{https://doi.or/10........ DOI placeholder (WILL BE FILLED IN BY TIB Open Publishing)}\\\medskip
        \footnotesize{\textcopyright\ Authors. This work is licensed under a \href{https://creativecommons.org/licenses/by/4.0/}{Creative Commons Attribution 4.0 International License}}\\\medskip
        \footnotesize{Published: (WILL BE FILLED IN BY TIB Open Publishing)}\\
    \end{flushleft}
    \begin{center}
        \LARGE\textbf{\contribtitle}\\\smallskip
        \Large{\contribsubtitle}
    \end{center}
    \normalsize{\firstauthor, \secondauthor, and \thirdauthor}
    \begin{center}
        \textsuperscript{1}\footnotesize{\firstauthoruniv}\\\medskip
        \textsuperscript{2}\footnotesize{\secondauthoruniv}\\\medskip
        \textsuperscript{3}\footnotesize{\thirdauthoruniv}\\\medskip
    \end{center}}
\makeatother

%===========[ TITLES ]==========================================================
%Chnage to your own information
%\newcommand{\contribtitle}{FAIR Knowledge Graphs for the Management of Health Information Systems}
\newcommand{\contribtitle}{SNIK Graph}
\newcommand{\contribsubtitle}{Visualizing Knowledge about Management of Hospital Information Systems}
%\newcommand{\contribsubtitle}{Graph-Based Exploration of an Hospital Information Systems}
\newcommand{\firstauthor}{Konrad Höffner\textsuperscript{1[https://orcid.org/0000-0001-7358-3217]}}
\newcommand{\firstauthoruniv}{Institute for Medical Informatics, Statistics and Epidemiology (IMISE), Germany}
\newcommand{\secondauthor}{Thomas Pause\textsuperscript{2[https://orcid.org/0000-0001-5832-4890]}}
\newcommand{\secondauthoruniv}{Institute for Medical Informatics, Statistics and Epidemiology (IMISE), Germany}
\newcommand{\thirdauthor}{Franziska Jahn\textsuperscript{3[https://orcid.org/0000-0002-7687-8544]}}%\todo{fj: orcid bitte anpassen}
\newcommand{\thirdauthoruniv}{Institute for Medical Informatics, Statistics and Epidemiology (IMISE), Germany}

%===========[ CONTENT ]=========================================================
\begin{document}
\maketitle

\paragraph{Abstract.}% The abstract should summarize the contents of the paper in short terms, i.e. in up to 250 words.
% ~ 150 words
Medical and health informatics integrates knowledge of business information systems, computer science and medicine.
As a comparatively young research discipline, it lacks a uniform terminology, especially for describing health information systems and their management.
SNIK, the semantic network of information management in hospitals, is an ontology that combines the knowledge of three textbooks, the IT4IT standard and an interview with a hospital CIO.
Graph-based visualizations allow teachers to intuitively convey relationships between a selection of those concepts and also allow students to explore the domain on their own.
We present SNIK Graph, a web-based client-side open-source Linked Data visualization for both classes and instances that is optimized for teaching and studying knowledge about the management of health information systems.
Due to the large amount of concepts and relationships, visualizing SNIK as a graph causes overplotting, which we solve using filters, searches and exploratory operations. 
%Students of medical informatics need to internalize complex relations between large amounts of concepts of health information systems (HIS) and their management.
%We present SNIK Graph, a graph-based linked data visualization web application that supports teaching textbook knowledge about HIS management.
\paragraph{Keywords:} Linked open data visualization, health information systems, information management,

\section*{Introduction}%
Medical and health informatics integrates knowledge of business information systems, computer science and medicine.
As a comparatively young research discipline, it lacks a uniform terminology, especially for describing health information systems and their management.
Several textbooks provide different perspectives of the discipline but the linear struture inherent in a book does not intuitively convey the highly connected nature of the concepts of the domain.
Thus, preparing students for practical or scientific work in the domain of health information systems requires an integrated system of concepts linking the terms of the different disciplines.
SNIK, the semantic network of information management in hospitals, is an ontology linking the knowledge of three textbooks~\cite{bb,ob,he}, the IT4IT standard~\cite{it4it}, and a CIO's knowledge about information management at a certain German university hospital.
It thus integrates theoretical knowledge of medical informatics and information systems as well as practical knowledge from industry and a health institutions.
%The SNIK ontology is visualized by SNIK Graph.

@@

%\subsection*{Motivation}%
%They also allow students of medical informatics internalize complex relations between large amounts of concepts of health information systems and their management~\cite{visualizationoflargeontologies}.
%\cite{snikgraphposter}
%
%A large amount of relevant knowledge is freely available as Linked Open Data.
%This structured approach is well suited for consumption by machines but basic access mechanisms like the text-based RDF serializations and SPARQL queries are not well suited for humans for large knowledge bases, especially for non-expert users.
%While the Semantic Web field has been providing a wealth of knowledge and is increasingly adopted by mainstream industries, not all use cases are yet supported by easy-to-use tools~\cite{semanticwebfield}.
%One such use case is graph-based exploration, where RDF data is displayed as a directed graph, with the resources as nodes and the triples as edges.
%As existing approaches \iffalse(see \cref{sec:relatedwork})\fi do not satisfy the needs of the students and researchers for exploring Hospital Information Management as Linked Open Data in the SNIK ontology~\cite{sniktec}, we implement \emph{SNIK Graph} as an easy-to-use graph-based web application, which we present in the following.
%To minimize duplicated 
%One such area is exploration, where users intuitively 
%Easy-to-use and freely available tools exist for 
%However, there is a lack of easy-to-use tools along the whole lifecycle of data~\cite{semanticwebfield}.
% - software is hard to publish
%The SNIK ontology~\cite{sniktec} contains knowledge about the management of hospital information systems.

@@

% version 3.18.0
\label{sec:relatedwork}
\section*{Related Work and Use Cases}
\citeauthor{linkeddatavisualization} \cite{linkeddatavisualization} describe different techniques and tools for Linked Data visualization, including graph-based visualization.
They identify 15 use cases and categorize them as related to T-Boxes, A-Boxes and both.
When showing instances, each RDF triple maps to an edge between two nodes representing its subject and object.
When visualizing ontologies, edges between class nodes can represent connections such as domain-range pairs and OWL restrictions that may be modelled using multiple triples.
SNIK Graph supports the following use cases:

@@

\section*{Preliminaries}
SNIK is created by manually extracting knowledge from sources such as textbooks.
It describes abstract concepts of hospital information systems and not any one in particular\footnote{With the exception of the CIO interview modelled as \url{http://www.snik.eu/ontology/ciox}.}.
Thus, SNIK is an ontology and models concepts as classes, such as \aurl{ob}{Consultant}.
However, the large number of classes and the reliance on RDF/RDFS formalisms has many characteristics of a knowledge base describing instances instead of classes.
This is a different approach than for example DBpedia, where many instances, such as \emph{Elephant}, actually represent abstract concepts with sets of individuals (classes).
Because of this modelling choice, the sources of SNIK are represented by a different subontologies that follow a meta ontology, the SNIK meta model, see ~\cref{fig:metamodel}.
Under the perspective of the meta model, the only parts of an abstract hospital information system that shall be described are roles (who), functions (does what) and entity types (which information is therefore needed).

@@

\subsection*{Design Goals and Requirements}
Our main goal is to visualize knowledge extracted from text books to users that may not have any Semantic Web experience.
The time and cognitive load of the users required to install, learn and operate the application should be as low as possible, so that they can focus on the data at hand and experience a benefit compared to only reading the textbook, such as when studying for an exam.
A web application is easiest to set up for the user, as it is operating system agnostic and does not need to be installed.
%\todo{Vorschlag von Michelle: Verwendung in der Lehre setzt intuitive/leicht erlernbare Benutzung des Graphen voraus um zu vermeiden wertvolle Vorlesungszeit an ein Tutorial zu verlieren --> Bekannet Tastaturshortcuts, Drag and Drop, übersichtliches }

\section*{Implementation and Setup}
As none of the investigated existing approaches support all our use cases and design goals and provide adequate performance on more than 4000 nodes, we implemented SNIK Graph as a web application using JavaScript (EcmaScript 2015) based on the graph visualization library Cytoscape.js~\cite{cytoscape}.
SNIK Graph is freely available as open source software\footnote{\url{https://github.com/imise/snik-cytoscape.js}}.
As a client-side tool\footnotemark{}, SNIK Graph can be adapted for other knowledge bases and ontologies and published on any HTML web server by cloning the repository, installing the NPM dependencies and adapting the configuration file settings including the SPARQL endpoint URL.
\footnotetext{Aside from a SPARQL endpoint.}%
An installation containing SNIK is published at \url{https://www.snik.eu/graph}.

\section*{Features}

\subsection*{Tabs and Containers}
\begin{figure}[h]
    \centering
    \includegraphics[width=\linewidth]{tabs.png}
    \caption{A session containing the default view and freely partitioned subgraphs containers.}
	\label{fig:tabs}
\end{figure}
\vspace{-3pt}
%For this subgraphs of the ontology have to be extracted and organized.
%A feature of SNIK Graph, tabs, allows multi-instance usage.
%One major issue in working with SNIK Graph was the missing ability of organizing teaching scenarios.
To organize teaching scenarios, independent SNIK Graph instances are loaded into containers\footnotemark{} that can be organized around the browser window.
\footnotetext{Implemented using the JavaScript GoldenLayout framework (\url{https://golden-layout.com/}).} 
These containers can either be accessed as tabs or freely partitioned see \cref{fig:tabs}.
The content and of the all containers can be saved and loaded as a session to continue developing a teaching scenario at a later point.

@@

\subsection*{Compound Layout}
\begin{figure}[h]
    \centering
    \includegraphics[width=\linewidth]{combine.pdf}
    \caption{Grouping of equivalent concepts from different sources in SNIK Graph (color coded).}% using the \enquote{combine matches} option.}
	\label{fig:combine}
\end{figure}
\vspace{-3pt}
%\noindent For citations of references, we prefer the use of square brackets and consecutive numbers, e.g. as shown by Author et al. [2], [3, pp. 5–10], as mentioned earlier [1], [3], [9]. The following bibliography provides the basic formats as a reference list with entries for journal articles [1], book chapter [2], as well as a URL [5]. For further guidance please refer to the \href{https://ieeeauthorcenter.ieee.org/wp-content/uploads/IEEE-Reference-Guide.pdf}{IEEE-Reference-Guide}.

@@

\subsection*{Role Use}
\iffalse
\todo{konrad an alle: was mache ich mit der sparql query? rausschmeissen, drin lassen oder umbauen? TP: drinlassen, aber direkt ans Bild hängen.. Sowas wie aus der Query folgt dieses Bild. Ich finde sowas sollte schon rein ins Paper}
\begin{figure}[h]
\begin{lstlisting}
SELECT DISTINCT ?inner ?middle ?outer ?outerx {
 <role> (rdfs:subClassOf|skos:closeMatch|^skos:closeMatch)* ?inner.
 ?inner meta:subTopClass meta:Role.
 OPTIONAL {
  ?inner ?p ?middle.
  ?middle meta:subTopClass meta:Function.
  OPTIONAL {
   ?middle ?q ?outer.
   ?outer meta:subTopClass meta:EntityType.
   OPTIONAL {?outer (skos:closeMatch|^skos:closeMatch|^rdfs:subClassOf)+ ?outerx.}
}}}
\end{lstlisting}
\caption{SPARQL query to generate the \emph{role use} visualization.}
\label{lst:roleuse}
\end{figure}
\vspace{-3pt}
\fi
A frequent question is, what a given role does and which information is needed for those functions represented by the entity types connected to those functions.
This question is visually answered by the \enquote{role use} feature, which arranges roles, functions and entity types in concentric circles, see \Cref{fig:classuse}:

@@

%\section*{Approach}
%Due to its large size, visualizing all of SNIK at once  overplotting.
%At the core of SNIK Graph is incremental exploration, which starts at one or more nodes and gradually expands them to related ones.

\subsection*{Search}
\begin{figure}[h!]
    \centering
    \includegraphics[width=\linewidth]{search.pdf}
    \caption{Visual search results for \enquote{Chief Information Officer} in a part of SNIK.}\label{fig:search}
\end{figure}

@@

\subsection*{Shortest Paths}
\begin{figure}[h!]
    \centering
    \includegraphics[width=\linewidth]{path.png}
    \caption{The shortest non-trivial path between \aurl{bb}{PatientIdentificationNumber} and \aurl{bb}{QualityOfData} in SNIK Graph.}
	\label{fig:path}
\end{figure}
\vspace{-3pt}
The shortest path is the most basic method of visualizing the relationship between two resources, see \cref{fig:path}.
We treat SNIK as an undirected graph for the calculation because the direction is often arbitrary.
Roles can be nested using \aurl{meta}{roleComponent} (\emph{has part}), which could have just as well been modelled as a \emph{part of} relation with reversed subject and object.
Another reason is that a set of resources that forms a graph connected by some symmetric property implies a complete subgraph, which negatively effects speed, memory and visibility of SNIK Graph.
For this reason, \aurl{skos}{closeMatch} and other symmetrical properties are often only sparsely modelled, and the other triples are implied.
As SPARQL endpoints, such as Virtuoso SPARQL employed by SNIK, do not perform reasoning to infer such implied triples such as those generated by symmetric properties, this would prevent the shortest path from including resources where only the opposite direction is explicitly modelled.%
\begin{figure}[h]
    \centering
    \includegraphics[width=0.5\linewidth]{hierarchy.pdf}
    \caption{Excerpt of the SNIK class hierarchy. Source: \cite{snikposter}.}
	\label{fig:hierarchy}
\end{figure}
\vspace{-3pt}
We also don't explicitly store triples inferred by transitive properties such as \aurl{rdfs}{subClassOf}, which prevents paths involving such properties from being squashed, such as the path between \emph{Project Manager} and \emph{HIS Stakeholder} in \cref{fig:hierarchy}.

@@

\subsection*{Neighbourhood Operations}
Exploration using neighbours, that is the successive uncovering of nodes adjacent to a starting node given by a user, is a common feature of tools such as \cite{lodlive} and \cite{vizlod}.%
\subsection*{Star}
The directed and undirected star operations show nodes connected to selected nodes via all paths that contain at most one property other than \aurl{skos}{closeMatch}.
The \emph{circle star} also rearranges the nodes using the force-directed layout locally on the currently visible subgraph.
\Cref{fig:star} shows a mind map of strategic information management, created by an undirected star, which can be used by a teacher to prepare a lecture about the topic.
\begin{figure}[h!]
    \centering
    \includegraphics[width=0.9\linewidth]{strategic-im.pdf}
    \caption{Undirected \emph{Star} of \aurl{bb}{StrategicInformationManagement}.}
	\label{fig:star}
\end{figure}
\vspace{-3pt}

%\todo{Konrad: Nimm den Star um "ob/ProjectManagement".} % ist inkludiert in \cref{fig:tabs}, refer?
The concept \emph{project management} both exists in a textbook~\cite{ob} and in the knowledge about information management in a German university hospital.
Grouping both concepts with the help of the star function the user detects that the terms have a different meaning according to their neighboorhood terms.
Whereas the text book deals with the management of single projects, in the CIO's world \enquote{project management} means managing multiple projects, as shown in the lower right container in \cref{fig:tabs}.

\subsection*{Mixed Operations}
\subsubsection*{Spiderworm}
A \emph{spiderworm} is a path from node \emph{A} to node \emph{B} combined with a \emph{star} of \emph{B}.
\Cref{fig:spiderworm} shows how we use a spiderworm to teach a student how the new concept \enquote{quality of data} is connected the the already introduced concept \enquote{patient identification number.}

@@

\section*{Conclusions and Future Work}
With SNIK Graph, we implement an adaptable open-source client-side graph-based Linked Data visualization that supports both A-Boxes and T-Boxes covering 11 of the 15 use cases presented in~\cite{linkeddatavisualization}.
We publish an installation visualizing the SNIK ontology and use it for teaching and self-learning of HIS management.
SNIK Graph is easy to setup and can be configured to visualize other knowledge bases and ontologies.  
Our main focus on improvement is the performance, which is mostly adequate on more than 4000 nodes, but suffers in two key areas:
% A problem is that related approaches are not available as open source ...
% this hinders analysis and slows progress, as effort is not shared and ...
%\section*{Evaluation}
%SNIK Graph is evaluated on the SNIK dataset, which contains textbook knowledge about the management of hospital information systems.

@@

%\subsection*{Future Work}
%\subsection*{Performance}
%On graphs of several thousand nodes,

>>>>

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{tabularx}
\usepackage{xcolor}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
%\usepackage{fancyhdr}
%\usepackage{titlesec}
%\usepackage{enumitem}
\usepackage{todonotes}
\usepackage{csquotes}
\usepackage{listings}
\lstset{language=SPARQL,breaklines=true}
\usepackage{siunitx}
\usepackage[labelsep=period]{caption}
\usepackage{mwe}
\usepackage{xfrac}
\usepackage[colorlinks=true,urlcolor=blue]{hyperref}
\usepackage{cleveref}
\usepackage{aurl}
\daurl{bb}{http://www.snik.eu/ontology/bb/}
\daurl{ob}{http://www.snik.eu/ontology/bb/}
\daurl{meta}{http://www.snik.eu/ontology/meta/}
\graphicspath{{./img/}}
%\addbibresource{paper.bib}
%\addbibresource{snik.bib}
%\addbibresource{relatedwork.bib}

\begin{document}
\title{SNIK Graph---Visualizing Knowledge about Management of Hospital Information Systems}
\thanks{This work is supported by the DFG (German Research Foundation) under the Project SNIK, Grant no. 1605/7-1 and 1387/8-1.}
% SNIK Graph is presented in poster form at \cite{snikgraphposter}.
\author{\IEEEauthorblockN{Konrad Höffner}
\IEEEauthorblockA{\textit{Institute for Medical Informatics, Statistics and Epidemiology} \\
\textit{Leipzig University}\\
Leipzig, Germany \\
\url{https://orcid.org/0000-0001-7358-3217}}
\and
\IEEEauthorblockN{Thomas Pause}
\IEEEauthorblockA{\textit{Institute for Medical Informatics, Statistics and Epidemiology} \\
\textit{Leipzig University}\\
Leipzig, Germany \\
\url{https://orcid.org/0000-0001-5832-4890}}
\and
\IEEEauthorblockN{Franziska Jahn}
\IEEEauthorblockA{\textit{Institute for Medical Informatics, Statistics and Epidemiology} \\
\textit{Leipzig University}\\
Leipzig, Germany \\
\url{https://orcid.org/0000-0002-7687-8544}}
\and
\and
\IEEEauthorblockN{Alfred Winter}
\IEEEauthorblockA{\textit{Institute for Medical Informatics, Statistics and Epidemiology} \\
\textit{Leipzig University}\\
Leipzig, Germany \\
%alfred.winter@imise.uni-leipzig.de}
\url{https://orcid.org/0000-0003-0179-954X}}
}
\maketitle

\begin{abstract}
% ~ 150 words
Medical and health informatics integrates knowledge of business information systems, computer science and medicine.
As a comparatively young research discipline, it lacks a uniform terminology, especially for describing health information systems and their management.
SNIK, the semantic network of information management in hospitals, is an ontology that combines the knowledge of three textbooks, the IT4IT standard and an interview with a hospital CIO.
Graph-based visualizations allow teachers to intuitively convey relationships between a selection of those concepts and also allow students to explore the domain on their own.
We present SNIK Graph, a web-based client-side open-source Linked Data visualization for both classes and instances that is optimized for teaching and studying knowledge about the management of health information systems.
Due to the large amount of concepts and relationships, visualizing SNIK as a graph causes overplotting, which we solve using filters, searches and exploratory operations. 
%Students of medical informatics need to internalize complex relations between large amounts of concepts of health information systems (HIS) and their management.
%We present SNIK Graph, a graph-based linked data visualization web application that supports teaching textbook knowledge about HIS management.
\end{abstract}
\begin{IEEEkeywords}
linked open data visualization,health information systems,information management
\end{IEEEkeywords}

\section{Introduction}%
Medical and health informatics integrates knowledge of business information systems, computer science and medicine.
As a comparatively young research discipline, it lacks a uniform terminology, especially for describing health information systems and their management.
Several textbooks provide different perspectives of the discipline but the linear struture inherent in a book does not intuitively convey the highly connected nature of the concepts of the domain.
Thus, preparing students for practical or scientific work in the domain of health information systems requires an integrated system of concepts linking the terms of the different disciplines.
SNIK, the semantic network of information management in hospitals, is an ontology linking the knowledge of three textbooks~\cite{bb,ob,he}, the IT4IT standard~\cite{it4it}, and a CIO's knowledge about information management at a certain German university hospital.
It thus integrates theoretical knowledge of medical informatics and information systems as well as practical knowledge from industry and a health institutions.
%The SNIK ontology is visualized by SNIK Graph.

@@

%\subsection{Motivation}%
%They also allow students of medical informatics internalize complex relations between large amounts of concepts of health information systems and their management~\cite{visualizationoflargeontologies}.
%\cite{snikgraphposter}
%
%A large amount of relevant knowledge is freely available as Linked Open Data.
%This structured approach is well suited for consumption by machines but basic access mechanisms like the text-based RDF serializations and SPARQL queries are not well suited for humans for large knowledge bases, especially for non-expert users.
%While the Semantic Web field has been providing a wealth of knowledge and is increasingly adopted by mainstream industries, not all use cases are yet supported by easy-to-use tools~\cite{semanticwebfield}.
%One such use case is graph-based exploration, where RDF data is displayed as a directed graph, with the resources as nodes and the triples as edges.
%As existing approaches \iffalse(see \cref{sec:relatedwork})\fi do not satisfy the needs of the students and researchers for exploring Hospital Information Management as Linked Open Data in the SNIK ontology~\cite{sniktec}, we implement \emph{SNIK Graph} as an easy-to-use graph-based web application, which we present in the following.
%To minimize duplicated 
%One such area is exploration, where users intuitively 
%Easy-to-use and freely available tools exist for 
%However, there is a lack of easy-to-use tools along the whole lifecycle of data~\cite{semanticwebfield}.
% - software is hard to publish
%The SNIK ontology~\cite{sniktec} contains knowledge about the management of hospital information systems.

@@

% version 3.18.0
\label{sec:relatedwork}
\section{Related Work and Use Cases}
%\citeauthor{linkeddatavisualization} \cite{linkeddatavisualization} describe different techniques and tools for Linked Data visualization, including graph-based visualization.
Po et al.~\cite{linkeddatavisualization} describe different techniques and tools for Linked Data visualization, including graph-based visualization.
They identify 15 use cases and categorize them as related to T-Boxes, A-Boxes and both.
When showing instances, each RDF triple maps to an edge between two nodes representing its subject and object.
When visualizing ontologies, edges between class nodes can represent connections such as domain-range pairs and OWL restrictions that may be modelled using multiple triples.
SNIK Graph supports the following use cases:

@@

\section{Preliminaries}
SNIK is created by manually extracting knowledge from sources such as textbooks.
It describes abstract concepts of hospital information systems and not any one in particular\footnote{With the exception of the CIO interview modelled as \url{http://www.snik.eu/ontology/ciox}.}.
Thus, SNIK is an ontology and models concepts as classes, such as \aurl{ob}{Consultant}.
However, the large number of classes and the reliance on RDF/RDFS formalisms has many characteristics of a knowledge base describing instances instead of classes.
This is a different approach than for example DBpedia, where many instances, such as \emph{Elephant}, actually represent abstract concepts with sets of individuals (classes).
Because of this modelling choice, the sources of SNIK are represented by a different subontologies that follow a meta ontology, the SNIK meta model, see ~\cref{fig:metamodel}.
Under the perspective of the meta model, the only parts of an abstract hospital information system that shall be described are roles (who), functions (does what) and entity types (which information is therefore needed).

@@

\subsection{Design Goals and Requirements}
Our main goal is to visualize knowledge extracted from text books to users that may not have any Semantic Web experience.
The time and cognitive load of the users required to install, learn and operate the application should be as low as possible, so that they can focus on the data at hand and experience a benefit compared to only reading the textbook, such as when studying for an exam.
A web application is easiest to set up for the user, as it is operating system agnostic and does not need to be installed.
%\todo{Vorschlag von Michelle: Verwendung in der Lehre setzt intuitive/leicht erlernbare Benutzung des Graphen voraus um zu vermeiden wertvolle Vorlesungszeit an ein Tutorial zu verlieren --> Bekannet Tastaturshortcuts, Drag and Drop, übersichtliches }

\section{Implementation and Setup}
As none of the investigated existing approaches support all our use cases and design goals and provide adequate performance on more than 4000 nodes, we implemented SNIK Graph as a web application using JavaScript (EcmaScript 2015) based on the graph visualization library Cytoscape.js~\cite{cytoscape}.
SNIK Graph is freely available as open source software\footnote{\url{https://github.com/imise/snik-cytoscape.js}}.
As a client-side tool\footnotemark{}, SNIK Graph can be adapted for other knowledge bases and ontologies and published on any HTML web server by cloning the repository, installing the NPM dependencies and adapting the configuration file settings including the SPARQL endpoint URL.
\footnotetext{Aside from a SPARQL endpoint.}%
An installation containing SNIK is published at \url{https://www.snik.eu/graph}.

\section{Features}

\subsection{Tabs and Containers}
\begin{figure}[h]
    \centering
    \includegraphics[width=\linewidth]{tabs.png}
    \caption{A session containing the default view and freely partitioned subgraphs containers.}
	\label{fig:tabs}
\end{figure}
\vspace{-3pt}
%For this subgraphs of the ontology have to be extracted and organized.
%A feature of SNIK Graph, tabs, allows multi-instance usage.
%One major issue in working with SNIK Graph was the missing ability of organizing teaching scenarios.
To organize teaching scenarios, independent SNIK Graph instances are loaded into containers\footnotemark{} that can be organized around the browser window.
\footnotetext{Implemented using the JavaScript GoldenLayout framework (\url{https://golden-layout.com/}).} 
These containers can either be accessed as tabs or freely partitioned see \cref{fig:tabs}.
The content and of the all containers can be saved and loaded as a session to continue developing a teaching scenario at a later point.

@@

\subsection{Compound Layout}
\begin{figure}[h]
    \centering
    \includegraphics[width=\linewidth]{combine.pdf}
    \caption{Grouping of equivalent concepts from different sources in SNIK Graph (color coded).}% using the \enquote{combine matches} option.}
	\label{fig:combine}
\end{figure}
\vspace{-3pt}
%\noindent For citations of references, we prefer the use of square brackets and consecutive numbers, e.g. as shown by Author et al. [2], [3, pp. 5–10], as mentioned earlier [1], [3], [9]. The following bibliography provides the basic formats as a reference list with entries for journal articles [1], book chapter [2], as well as a URL [5]. For further guidance please refer to the \href{https://ieeeauthorcenter.ieee.org/wp-content/uploads/IEEE-Reference-Guide.pdf}{IEEE-Reference-Guide}.

@@

\subsection{Role Use}
\iffalse
\todo{konrad an alle: was mache ich mit der sparql query? rausschmeissen, drin lassen oder umbauen? TP: drinlassen, aber direkt ans Bild hängen.. Sowas wie aus der Query folgt dieses Bild. Ich finde sowas sollte schon rein ins Paper}
\begin{figure}[h]
\begin{lstlisting}
SELECT DISTINCT ?inner ?middle ?outer ?outerx {
 <role> (rdfs:subClassOf|skos:closeMatch|^skos:closeMatch)* ?inner.
 ?inner meta:subTopClass meta:Role.
 OPTIONAL {
  ?inner ?p ?middle.
  ?middle meta:subTopClass meta:Function.
  OPTIONAL {
   ?middle ?q ?outer.
   ?outer meta:subTopClass meta:EntityType.
   OPTIONAL {?outer (skos:closeMatch|^skos:closeMatch|^rdfs:subClassOf)+ ?outerx.}
}}}
\end{lstlisting}
\caption{SPARQL query to generate the \emph{role use} visualization.}
\label{lst:roleuse}
\end{figure}
\vspace{-3pt}
\fi
A frequent question is, what a given role does and which information is needed for those functions represented by the entity types connected to those functions.
This question is visually answered by the \enquote{role use} feature, which arranges roles, functions and entity types in concentric circles, see \Cref{fig:classuse}:

@@

%\section{Approach}
%Due to its large size, visualizing all of SNIK at once  overplotting.
%At the core of SNIK Graph is incremental exploration, which starts at one or more nodes and gradually expands them to related ones.

\subsection{Search}
\begin{figure}[h!]
    \centering
    \includegraphics[width=\linewidth]{search.pdf}
    \caption{Visual search results for \enquote{Chief Information Officer} in a part of SNIK.}\label{fig:search}
\end{figure}

@@

\subsection{Shortest Paths}
\begin{figure}[h!]
    \centering
    \includegraphics[width=\linewidth]{path.png}
    \caption{The shortest non-trivial path between \aurl{bb}{PatientIdentificationNumber} and \aurl{bb}{QualityOfData} in SNIK Graph.}
	\label{fig:path}
\end{figure}
\vspace{-3pt}
The shortest path is the most basic method of visualizing the relationship between two resources, see \cref{fig:path}.
We treat SNIK as an undirected graph for the calculation because the direction is often arbitrary.
Roles can be nested using \aurl{meta}{roleComponent} (\emph{has part}), which could have just as well been modelled as a \emph{part of} relation with reversed subject and object.
Another reason is that a set of resources that forms a graph connected by some symmetric property implies a complete subgraph, which negatively effects speed, memory and visibility of SNIK Graph.
For this reason, \aurl{skos}{closeMatch} and other symmetrical properties are often only sparsely modelled, and the other triples are implied.
As SPARQL endpoints, such as Virtuoso SPARQL employed by SNIK, do not perform reasoning to infer such implied triples such as those generated by symmetric properties, this would prevent the shortest path from including resources where only the opposite direction is explicitly modelled.%
\begin{figure}[h]
    \centering
    \includegraphics[width=0.5\linewidth]{hierarchy.pdf}
    \caption{Excerpt of the SNIK class hierarchy. Source: \cite{snikposter}.}
	\label{fig:hierarchy}
\end{figure}
\vspace{-3pt}
We also don't explicitly store triples inferred by transitive properties such as \aurl{rdfs}{subClassOf}, which prevents paths involving such properties from being squashed, such as the path between \emph{Project Manager} and \emph{HIS Stakeholder} in \cref{fig:hierarchy}.

@@

\subsection{Neighbourhood Operations}
Exploration using neighbours, that is the successive uncovering of nodes adjacent to a starting node given by a user, is a common feature of tools such as \cite{lodlive} and \cite{vizlod}.%
\subsection{Star}
The directed and undirected star operations show nodes connected to selected nodes via all paths that contain at most one property other than \aurl{skos}{closeMatch}.
The \emph{circle star} also rearranges the nodes using the force-directed layout locally on the currently visible subgraph.
\Cref{fig:star} shows a mind map of strategic information management, created by an undirected star, which can be used by a teacher to prepare a lecture about the topic.
\begin{figure}[h!]
    \centering
    \includegraphics[width=0.9\linewidth]{strategic-im.pdf}
    \caption{Undirected \emph{Star} of \aurl{bb}{StrategicInformationManagement}.}
	\label{fig:star}
\end{figure}
\vspace{-3pt}

%\todo{Konrad: Nimm den Star um "ob/ProjectManagement".} % ist inkludiert in \cref{fig:tabs}, refer?
The concept \emph{project management} both exists in a textbook~\cite{ob} and in the knowledge about information management in a German university hospital.
Grouping both concepts with the help of the star function the user detects that the terms have a different meaning according to their neighboorhood terms.
Whereas the text book deals with the management of single projects, in the CIO's world \enquote{project management} means managing multiple projects, as shown in the lower right container in \cref{fig:tabs}.
As all neighbourhood and path operations automatically select all resulting nodes, repeated application of a star without clearing the selection results in an incremental uncovering of the whole connected graph around the starting node.

\subsection{Mixed Operations}
\subsubsection{Spiderworm}
A \emph{spiderworm} is a path from node \emph{A} to node \emph{B} combined with a \emph{star} of \emph{B}.
\Cref{fig:spiderworm} shows how we use a spiderworm to teach a student how the new concept \enquote{quality of data} is connected the the already introduced concept \enquote{patient identification number.}

@@

\section{Conclusions and Future Work}
With SNIK Graph, we implement an adaptable open-source client-side graph-based Linked Data visualization that supports both A-Boxes and T-Boxes covering 11 of the 15 use cases presented in~\cite{linkeddatavisualization}.
We publish an installation visualizing the SNIK ontology and use it for teaching and self-learning of HIS management.
SNIK Graph is easy to setup and can be configured to visualize other knowledge bases and ontologies.  
Our main focus on improvement is the performance, which is mostly adequate on more than 4000 nodes, but suffers in two key areas:
% A problem is that related approaches are not available as open source ...
% this hinders analysis and slows progress, as effort is not shared and ...
%\section{Evaluation}
%SNIK Graph is evaluated on the SNIK dataset, which contains textbook knowledge about the management of hospital information systems.

@@

%\subsection{Future Work}
%\subsection{Performance}
%On graphs of several thousand nodes,
